# Tutorial for RBET

RBET (Reference-based batch effect testing) is designed to test the differences between two or more batches in single-cell data. RBET is implemented in a two-step strategy, which are (1) reference genes selection, and (2) batch effect testing. Reference genes can be selected either from candidate housekeeping genes or directly from data, which correspond to literature-based RBET and data-based RBET. In the following, we will give toy examples to both methods.

```{r load}
library(RBET)
```

## Literature-based RBET

When candidate housekeeping genes and cell type information are provided, we recommend to use literature-based RBET for higher credibility. Here is a toy example of pancreas dataset with two batches and two cell types. Please note that reference genes selection should be implemented on raw data.

```{r data1}
seu = readRDS("example_data.rds")
DefaultAssay(seu) = "RNA"
```

(Optional) If you have multiple layers in a single slot, join them together.

```{r opt1}
seu = JoinLayers(seu)
```

We provide lists of candidate genes for 23 tissues or cell types in `reference_genes`, which are housekeeping genes collected from previous literature. If a more reliable catalogue of housekeeping genes is available for your data, we suggest to utilize them as candidates.

```{r candidate1}
names(reference_genes)
can_genes = reference_genes[["Pancreas"]]
```

Then, we select reference genes from candidates.

```{r reference1}
ref_genes = rg_sel_literature(seu, batch = "batch", celltype = "celltype", rgene = can_genes)
ref_genes
```

Finally, we test batch effect on the integrated data.

```{r rbet1}
DefaultAssay(seu) = "integrated"
res = RBET(seu, ref_genes, batch = "batch")
res
```

## Data-based RBET

We also provide a data-driven RBET that operates independently of cell type information and candidate housekeeping genes. Here is a toy example of pancreas dataset with two batches. Please note that reference genes selection should be implemented on raw data.

```{r data2}
seu = readRDS("example_data.rds")
DefaultAssay(seu) = "RNA"
```

(Optional) If you have multiple layers in a single slot, join them together.

```{r opt2}
seu = JoinLayers(seu)
```

Then, we select reference genes from data. This step will take some time. If the number of reference genes is less than 2, we suggest to rerun this step with a larger `p`.

```{r reference2}
ref_genes = rg_sel_data(seu, batch = "batch", p = 0.05)
ref_genes
```

Finally, we test batch effect on the integrated data.

```{r rbet2}
DefaultAssay(seu) = "integrated"
res = RBET(seu, ref_genes, batch = "batch")
res
```
